---
title: "MD06 Demo, Part I"
format: html
editor: visual
---

Load packages and set theme.

```{r}
#| message: false
library(tidyverse)
library(moderndive)
library(ggthemes)
library(patchwork)
theme_set(theme_light())
```

Let's look at the MD evaluations data.

```{r}
data(evals)
glimpse(evals)
```

We are going to ask whether the "beauty" of an instructor predicts their teaching evaluations.

```{r}
d <- evals |>
  rename(bty = bty_avg,    # just shorter to type
         sex = gender)     # actually what they have

glimpse(d)
```

Let's look at the first few cases.

```{r}
head(d)
```

Here's the regression from last week.

```{r}
mod1 <- lm(score ~ bty,
           data = d)

get_regression_table(mod1)
```

Let's look at the predictions and residuals.

```{r}
mod1_preds <- get_regression_points(mod1)

head(mod1_preds)
```

Here's the regression line (blue).

```{r,echo=FALSE}
ggplot(d,
       aes(x = bty,
           y = score)) +
  geom_jitter(alpha = .3) +
  geom_smooth(method = "lm",
              se = FALSE) +
  labs(x = "Beauty",
       y = "Evaluation",
       title = "Simple regression results")
```

Here are the residuals.

```{r,echo=FALSE}
ggplot(mod1_preds,
       aes(x = bty,
           y = residual)) +
  geom_jitter(alpha = .3) +
  geom_hline(yintercept = 0,
             color = "blue") +
  labs(x = "Beauty",
       y = "Residual",
       title = "Simple regression residuals")
```

One way to quantify how well a predictor predicts the outcome is to use the **variance**. We've seen this already but let's review. Here's the formula. $\bar{y}$ is the mean of $y$.

$$
V(y) = \frac{1}{n-1}\sum_{i=1}^{n}(y_i - \bar{y})^2
$$

Since our outcome is evaluation score, we can just do that.

```{r}
var_y <- d |> 
  pull(score) |> 
  var()

var_y
```

This is equivalent to the variance of the "residuals" when we just guess the mean value for every person!

```{r, echo=FALSE}
ggplot(d,
       aes(x = bty,
           y = score)) +
  geom_jitter(alpha = .3) +
  geom_hline(yintercept = mean(d$score),
             color = "blue") +
  labs(x = "Beauty",
       y = "Evaluation",
       title = "Guessing the mean for everyone")
```

Adding `beauty` as a predictor improves our guess.

```{r, echo=FALSE}
ggplot(d,
       aes(x = bty,
           y = score)) +
  geom_jitter(alpha = .3) +
  geom_hline(yintercept = mean(d$score),
             color = "blue") +
  geom_smooth(method = "lm",
              se = FALSE,
              color = "red",
              linetype = "dashed") +
  labs(x = "Beauty",
       y = "Evaluation",
       title = "Mean vs. regression line")
```

Now let's see what the spread looks like if we look at the residuals from the regression line.

```{r}
var_yhat1 <- mod1_preds |> 
  pull(residual) |> 
  var()

var_yhat1
```

If we take the ratio of these and subtract it from one, that gives us the $R^2$ or "proportion of variance explained" by beauty scores.

```{r}
1 - (var_yhat1 / var_y)
```

It looks like "beauty" can account for about `r round((1 - (var_yhat1 / var_y)) * 100, 1)` percent of the variance in the evaluation scores.

In other words, if we try to guess instructors' evaluation scores, our guesses will be `r round((1 - (var_yhat1 / var_y)) * 100, 1)` percent _less wrong_ on average if we know the instructor's beauty rating. 

We can get this from `broom::glance()` or `moderndive::get_regression_summaries()` without doing it manually. The latter will be a more curated list of things you'll need for this course.

```{r}
broom::glance(mod1)
moderndive::get_regression_summaries(mod1)
```


```{r}
mod_alt <- lm(score ~ age,
           data = d)

get_regression_table(mod_alt)
get_regression_summaries(mod_alt)

mod_alt_preds <- get_regression_points(mod_alt)
```

```{r}
ggplot(d,
       aes(x = age,
           y = score)) +
  geom_jitter(alpha = .3) +
  geom_smooth(method = "lm",
              se = FALSE) +
  labs(x = "Age",
       y = "Evaluation")
```

```{r}
ggplot(mod_alt_preds,
       aes(x = age,
           y = residual)) +
  geom_jitter(alpha = .3) +
  geom_hline(yintercept = 0,
             color = "blue") +
  labs(x = "Age",
       y = "Residual")
```

```{r}
ggplot(d,
       aes(x = age,
           y = score)) +
  geom_jitter(alpha = .3) +
  geom_hline(yintercept = mean(d$score),
             color = "blue") +
  geom_smooth(method = "lm",
              se = FALSE,
              color = "red",
              linetype = "dashed") +
  labs(x = "Age",
       y = "Evaluation")
```

